Inference for Trans-dimensional Bayesian Models with Diffusive Nested Sampling

### 2014-08-28 Documents/Paper/ms.tex
Working on results

@@

For the interim prior for the total fluxes of the galaxies, and for the
widths of the galaxies, we used the Pareto
distribution. The probability density function for a variable $x$ given
hyperparameters $x_{\rm min}$ and $a$ is
\begin{eqnarray}
p(x | x_{\rm min}, a) &=&
\left\{
\begin{array}{lr}
\frac{a x_{\rm min}^{a}}{x^{a + 1}}, & x \geq x_{\rm min}\\
0, & \textnormal{otherwise.}
\end{array}
\right.
\end{eqnarray}
Since we are using this for both the fluxes and the widths of the galaxies,
there will be four hyperparameters in total: two lower limits (one for the
fluxes and one for the widths) and two slopes. As in Section~\ref{sec:sinewaves}
we allowed the noise standard deviation $\sigma$ to be a free parameter with
a log-uniform prior between $10^{-3}$ and $10^3$.

@@

@@

In general, it is possible to compute the number of components that have been
affected by the proposal. If this is greater than or equal to $N$, we compute
the mock signal from scratch, otherwise we subtract the effect of those
components that have been removed and add in the effect of those that have been
added. In the C++ implementation of this work, when a proposal takes
place, the difference between the old and new model is cached and is available
to the user, so that the likelihood can be updated rather than computed
from scratch.

>>>>

The marginal likelihood estimate returned by DNS for this data was
$\log(\mathcal{Z}) = -1410.8$, and the information, or Kullback-Leibler
divergence from the prior to the posterior, was estimated to be 29.7 nats.

@@

\subsection{Priors}
For the interim prior for the total fluxes of the galaxies, and for the
widths of the galaxies, we used the Pareto
distribution. The probability density function for a variable $x$ given
hyperparameters $x_{\rm min}$ and $a$ is
\begin{eqnarray}
p(x | x_{\rm min}, a) &=&
\left\{
\begin{array}{lr}
\frac{a x_{\rm min}^{a}}{x^{a + 1}}, & x \geq x_{\rm min}\\
0, & \textnormal{otherwise.}
\end{array}
\right.
\end{eqnarray}
Since we are using this for both the fluxes and the widths of the galaxies,
there will be four hyperparameters in total: two lower limits (one for the
fluxes and one for the widths) and two slopes. The lower limits were assigned
log-uniform priors between $10^{-3}$ and $10^3$, and uniform priors between 0
and 1 were assigned to the {\it inverses} of the slopes. This latter choice
was based on the fact that a Pareto distribution for $x$ is an exponential
distribution for $\ln(x)$, with scale length $1/a$.

The interim priors for the radius ratios $u$ and the flux ratios $w$ were both
uniform distributions. Since $u \in [0,1]$ and $v \in [0,1]$, the upper limits
of the uniform interim priors must be between 0 and 1. Therefore the prior
for the upper limit was a uniform distribution between 0 and 1, and the prior
for the lower limit given the upper limit was a uniform distribution between 0
and the upper limit.

As in Section~\ref{sec:sinewaves}
we allowed the noise standard deviation $\sigma$ to be a free parameter with
a log-uniform prior between $10^{-3}$ and $10^3$.

\subsection{Results}
The results from running DNS on the simulated galaxy data are shown in
Figures~\ref{fig:galaxyfield_likelihood} and~\ref{fig:N_galaxy_result}.
This problem, like the sinusoidal problem, exhibits a phase transition, which
can be seen in Figure~\ref{fig:galaxyfield_likelihood}. The phase transition
separates models which only fit the brightest galaxies from models that
also fit the faint galaxies close to the noise level. This phase transition
does not affect the posterior distribution, as the low-$N$ solutions are
eventually found to have very low importance.

@@

The marginal likelihood estimate returned by DNS for this data was
$\log(\mathcal{Z}) = -319707.6$, and the information, or Kullback-Leibler
divergence from the prior to the posterior, was estimated to be 550.2 nats.

@@

In general, it is possible to compute the number of components that have been
affected by the proposal. If this is greater than or equal to $N$, we compute
the mock signal from scratch, otherwise we subtract the effect of those
components that have been removed and add in the effect of those that have been
added. In the C++ implementation of this work, when a proposal takes
place, the difference between the old and new model is cached and is available
to the user, so that the likelihood can be updated rather than computed
from scratch. Using this technique offers a speed advantage of a factor of
$\sim 2$ for both of the example problems. The sinusoidal example could be
solved in a few minutes on a desktop computer, whereas the galaxy example took
much longer, about a day, to execute.
